A Kochen-Specker system has at least 22 vectors

### 2015-01-30 paper/main.tex
paper: remove watermark
Related work & typo's

\usepackage{draftwatermark}

@@

First, in Subsection~\ref{sec:ksgraphs},
we repeat a part of AOW's work, in particular the reduction of
KS systems to graphs.
The bottleneck of their search was the sheer number of graphs
and the deciding whether such graphs are embeddable.
In Section~\ref{sec:ilb},
we improve upon their reduction,
to cut down the number of graphs to consider drastically,
and state the results of our main computation.
Finally, in Section~\ref{sec:emb},
we describe our practical embeddability test.

The software and results of the various computations performed for
this paper, can be found here\cite{GH}.

@@

\begin{comp}\label{comp:types}
    We have enumerated all embeddable graphs of less than~$17$ vertices
    and determined a lower bound at which a particular $1$- or $2$-type occurs,
    omitting the trivial types~$\{(0),(1)\}$
    and~$\{(0,0),(0,1),(1,0),(1,1)\}$.%
        \footnote{Source code at \texttt{code/comp5.py} of \cite{GH}.}
    \begin{center}
    \begin{tabular}{ll}
        $1/2$-type & $\#G$ \\ \hline
        $\{(0,0), (1,0), (0,1)\}$ &
            non-trivially $\geq 10$\\ 
        $\{(0,0), (1,0), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1))\}$ & $\geq 15$ \\
        $\{(0,0), (1,0))\}$ & $\geq 15$ \\
        $\{(0)\}$ & $\geq 15$ \\
        $\{(0,1),(1,0)\}$ & $\geq 16$ \\
        other & $\geq 17$
    \end{tabular}
    \end{center}
    The type~$\{(0,0),(1,0),(0,1)\}$ occurs in the embeddable two-vertex
    graph
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
            (0,0) node[p]{} -- (0.3,0) node[p]{} ;
    \end{tikzpicture}.
    Because the two vertices are adjacent, this occurance of
    the type is called trivial.
\end{comp}

@@

            The type of~$(a_1,a_2)$ in~$A$ cannot contain~$(1,0)$
            or~$(0,1)$ for otherwise~$G$ can be colored.
            It also cannot contain~$(1,1)$ as~$a_1$ and~$a_2$
            are adjacent.
            Thus both~$a_1$ and~$a_2$ have fixed color~$0$ in~$A$.
            Hence~$\#A\geq 17$ by Computation~\ref{comp:types}.
            Consequently~$\#G \geq 32$. Contradiction
            with minimality.
        \end{enumerate}
    \item Suppose~$a_1 = a_2$ and~$b_1 \neq b_2$.
            This leads to a contradication in the same way
            as in case~\ref{tricase1}.
    \item Apparently~$a_1 \neq a_2$ and~$b_1 \neq b_2$.
        The type of~$(a_1,a_2)$ in~$A$ cannot contain~$(0,0)$,
        for otherwise~$G$ is colorable.
        Similarly, the type of~$(b_1,b_2)$ in~$B$ cannot contain~$(0,0)$.
        Thus both~$\#A \geq 17$ and~$\#B \geq 17$.
        Hence~$\#G \geq 34$. Contradiction with minimality. \qed
    \end{enumerate}
\end{myproof}

@@

we proceded as follows.

@@

If we only consider the points used to obtain the
contradiction, we find an unembeddable subgraph of the system
containing 13 vertices\footnote{These are the vetices
0,2,3,4,6,8,9,10,13,14,15,17 and 18.}.
In the appendix we included the  derivation of contradiction.
In theory, this approach can be mechanically performed by
a computer.  However, there are too many variables
to solve this in reasonable time.

@@

@@

Enumerating all candidate graphs of less than 31 vertices
is computationally infeasable.
To bridge the enormous the gap between 22 and 31,
requires a new insight.
Most likely, this would be a new
practical restriction on which graphs to consider.
In particular
(following Prop.~\ref{prop:ks-enoughtriangles})
a fast algorithm to enumerate
square-free graphs of minimal vertex order three
where each node is part of a triangle,
would probably allow to test for KS-systems
up to and including 24 vertices.

@@

\appendix
\clearpage
\section{Appendix}
In this appendix, we show by hand, that indeed the 19 vertex graph
considered at the start of section 3 %TODO
is indeed not embeddable.
We restrict the calculation to the~$13$-vertex unembeddable
subgraph.

@@

\begin{equation}
\frac{\sqrt{1-{x_3}^2}}{x_3} = \frac{-\sqrt{1-{x_3}^2}}{x_3 {x_1}^2  (1-{x_1}^2)}
\end{equation}
But then, eliminating the $x_3$ terms we find:
\begin{equation}
{x_1}^2 (1-{x_1}^2) = -1
\end{equation}
Which is our desired contradicton.

>>>>

@@

First, in Subsection~\ref{sec:ksgraphs},
we repeat a part of AOW's work, in particular the reduction of
KS systems to graphs.
The bottleneck of their search was the sheer number of graphs
and the deciding whether such graphs are embeddable.
In Section~\ref{sec:ilb},
we improve upon their reduction,
to cut down the number of graphs to consider drastically,
and state the results of our main computation.
Finally, in Section~\ref{sec:emb},
we describe our practical embeddability test.
In Section~\ref{sec:related},
we discuss related work.
The software and results of the various computations performed for
this paper, can be found here\cite{GH}.

This is an extended version of the paper published in proceedings
of the 11th workshop on Quantum Physics and Logic.\cite{UW14}
Compared to the results presented on the workshop,
we have improved the lower-bound to~$22$.
We also include new restrictions on the graph enumeration
of which we would like to emphasize~Prop.~\ref{prop:ks-enoughtriangles},
which promises to push the bound when implemented efficiently.
Furthermore we expand on methods to prove unembeddability.
Finally we discuss how the related results in the literature
compare to ours.

@@

\begin{comp}\label{comp:types}
    We have enumerated all embeddable graphs of less than~$17$ vertices
    and determined a lower bound at which a particular $1$- or $2$-type occurs,
    omitting the trivial types~$\{(0),(1)\}$
    and~$\{(0,0),(0,1),(1,0),(1,1)\}$.%
        \footnote{Source code at \texttt{code/comp5.py} of \cite{GH}.}
    \begin{center}
    \begin{tabular}{ll}
        $1/2$-type & $\#G$ \\ \hline
        $\{(0,0), (1,0), (0,1)\}$ &
            non-trivially $\geq 10$\\ 
        $\{(0,0), (1,0), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1), (1,1)\}$ & $\geq 10$ \\
        $\{(0,0), (0,1))\}$ & $\geq 15$ \\
        $\{(0,0), (1,0))\}$ & $\geq 15$ \\
        $\{(0)\}$ & $\geq 15$ \\
        $\{(0,1),(1,0)\}$ & $\geq 16$ \\
        other & $\geq 17$
    \end{tabular}
    \end{center}
    The type~$\{(0,0),(1,0),(0,1)\}$ occurs in the embeddable two-vertex
    graph
    \begin{tikzpicture}[thick,scale=1.4,
        p/.style={circle, draw, fill=white,
                        inner sep=0pt, minimum width=5pt}]
            \draw
            (0,0) node[p]{} -- (0.3,0) node[p]{} ;
    \end{tikzpicture}.
    Because the two vertices are adjacent, this occurrence of
    the type is called trivial.
\end{comp}

@@

            The type of~$(a_1,a_2)$ in~$A$ cannot contain~$(1,0)$
            or~$(0,1)$ for otherwise~$G$ can be colored.
            It also cannot contain~$(1,1)$ as~$a_1$ and~$a_2$
            are adjacent.
            Thus both~$a_1$ and~$a_2$ have fixed color~$0$ in~$A$.
            Hence~$\#A\geq 17$ by Computation~\ref{comp:types}.
            Consequently~$\#G \geq 32$. Contradiction
            with minimality.
        \end{enumerate}
    \item Suppose~$a_1 = a_2$ and~$b_1 \neq b_2$.
            This leads to a contradiction in the same way
            as in case~\ref{tricase1}.
    \item Apparently~$a_1 \neq a_2$ and~$b_1 \neq b_2$.
        The type of~$(a_1,a_2)$ in~$A$ cannot contain~$(0,0)$,
        for otherwise~$G$ is colorable.
        Similarly, the type of~$(b_1,b_2)$ in~$B$ cannot contain~$(0,0)$.
        Thus both~$\#A \geq 17$ and~$\#B \geq 17$.
        Hence~$\#G \geq 34$. Contradiction with minimality. \qed
    \end{enumerate}
\end{myproof}

@@

we proceeded as follows.

@@

If we only consider the points used to obtain the
contradiction, we find an unembeddable subgraph of the system
containing 13 vertices\footnote{These are the vertices
0,2,3,4,6,8,9,10,13,14,15,17 and 18.}.
In the appendix we included the  derivation of contradiction.
In theory, this approach can be mechanically performed by
a computer.  However, there are too many variables
to solve this in reasonable time.

@@

\section{Related work}
\label{sec:related}
This paper improves the lower bound of
Arends, Ouaknine and Wampler\cite{a09,aow11}
on the size of a minimal Kochen-Specker system.
Pavi\v{c}i\'c, Merlet, McKay and Norman\cite{pavivcic}
count the size of a Kochen-Specker system differently
and have demonstrated a different bound.
\begin{definition}
    A Kochen-Specker graph is called \keyword{complete}
        if every edge is part of a triangle.
\end{definition}
Pavi\v{c}i\'c et al have given a lower bound on the size
of the minimal \emph{complete} Kochen-Specker system.
Similar to our approach, they reduce the problem
to a question about discrete structures
which can be considered as a subclass of graphs.
They discovered several uncolorable candidates
of which they could prove unembeddability
using either symbolic computation or interval analysis.

Clearly, a bound on the minimal size of a KS system
is also a bound on the minimal size of a complete system.
Conversely, a KS system can be completed by adding missing vectors.
However, it is not clear how many vectors have to be added in general
to complete the system.
A history of the different bounds on 3-dimensional Kochen-Specker
systems is given in the table below.
\begin{center}
    \begin{tabular}{llll}
        Authors & year & KS & complete KS \\ \hline
        Kochen and Specker\cite{ks} & 1975 & $\leq 117$ & $\leq 192$ \\
        Penrose, Peres\cite{peres}
                (independendly)& 1991 & $\leq 33$ & $\leq 57$ \\
        Conway & $\sim$1995 & $\leq 31$ & $\leq 51$ \\
        UW\cite{UW14} & 2014 & $\geq 22$ & $\geq 22$ \\
        Pavi\v{c}i\'c et al.\cite{pavivcic}& 2004 & $?$ & $\geq 30$ \\
    \end{tabular}
\caption{A history of the bounds on the size of the minimal
    (complete) three-dimensional Kochen-Specker system.}
\end{center}

@@

Enumerating all candidate graphs of less than 31 vertices
is computationally infeasible.
To bridge the enormous the gap between 22 and 31,
requires a new insight.
Most likely, this would be a new
practical restriction on which graphs to consider.
In particular
(following Prop.~\ref{prop:ks-enoughtriangles})
a fast algorithm to enumerate
square-free graphs of minimal vertex order three
where each node is part of a triangle,
would probably allow to test for KS-systems
up to and including 24 vertices.

@@

\appendix
\clearpage
\section{Appendix}
In this appendix, we show by hand, that indeed the 19 vertex graph
considered at the start of section 3 %TODO
is not embeddable.
We restrict the calculation to the~$13$-vertex unembeddable
subgraph.

@@

\begin{equation}
\frac{\sqrt{1-{x_3}^2}}{x_3} = \frac{-\sqrt{1-{x_3}^2}}{x_3 {x_1}^2  (1-{x_1}^2)}
\end{equation}
But then, eliminating the $x_3$ terms we find:
\begin{equation}
{x_1}^2 (1-{x_1}^2) = -1
\end{equation}
Which is our desired contradiction.
